---
title: "Basic Models-Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Models-Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r}
rm(list=ls())
```

```{r setup}
library(blouch)
```


Here I walk you through the basic models of *Blouch*. By basic I mean either the direct effect, adaptive, or model with a combination of both types of predictors.

## Direct effect model
*Blouch* implements the model of constrained evolution (Hansen & Bartoszek, 2012) known as the direct effect model, previously implemented in Grabowski et al. (2016), which can be used to test for allometric constraints.

First we will create a phylogeny by randomly sampling from the 10K Trees phylogeny
```{r Create_phylogeny}
########################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label


```

First we will simulate X and Y data using a generative model for the direct effect model. We will use the treeplyr package (Uyeda and Harmon, 2014) make.treedata function to combine the data and tree based on the row names, which has the taxa names. See https://github.com/uyedaj/treeplyr for more on this package.

We will set the intercept (alpha term) to 2 and the slope to 0.25. The half-life will be set to 0.1 or 10% of tree height, with the stationary variance (Vy) at 0.01. We will use the *Blouch* helper function calc_direct_V to calculate the direct effect variance/covariance matrix, and the fastBM function from phytools (Revell, 2011) to simulate the X variable following a BM process.

```{r Direct_effect_model_setup}
########################################################################################################
#Direct Effect Model
########################################################################################################
#Setup parameters
Z_direct<-1 #Number of traits
hl<-0.1 #half life
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));

vX0<-0
vY0 <- 0

Sxx<-10 #Variance of BM process

V<-calc_direct_V(phy,sigma2_y,a) #Calculate V - variance/covariance matrix
X<-phytools::fastBM(phy,a=vX0,sig2=Sxx,internal=FALSE) #Simulate X BM variable on tree, with scaling 10
phytools::phenogram(phy,X,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data

```


Next we set the intercept and slope parameters, and simulate mu before using it and V to simulate Y. Finally we make a quick plot of the results.

```{r}
alpha<-2 #Intecept
beta<-0.25 #Slope

mu<-alpha+X*beta #Simulate mu for Y

Y<-MASS::mvrnorm(n=1,mu,V) #Simulate direct effect Y trait

df<-data.frame(Y=Y,X=X)
names(df)<-c("Y","X")

ggplot2::ggplot(data=df,ggplot2::aes(x=X,y=Y))+
  ggplot2::geom_point()
summary(lm(Y~X,df))

```

Next we will simulate measurement error - we will use a standard deviation of measurement error of 0.01, which we will provide to *Blouch* as a vector (X_error and Y_error), and use the rnorm function to add error to our X and Y variables. In other words, we are telling *Blouch* that the estimated error on X and Y is 0.01, and providing it with X and Y variables that are offset by a random amount of error with this standard deviation.
```{r}
X_error<-rep(0.01,N)
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-X+rnorm(N,0,0.01)

```

## Data setup for *Blouch*

Next we will use the treeplyr package (Uyeda and Harmon, 2014) make.treedata function to combine the data and tree based on the taxa names. See https://github.com/uyedaj/treeplyr for more on this package.
```{r}
############################################################################################################
#Make trdata file
trait.data<-data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error))
trdata<-treeplyr::make.treedata(phy,trait.data)

```

We will use the helper function blouch.direct.prep() to setup the dat file for Stan. Here the name of the column in trdata$dat that contains the response variable is "Y_with_error", the associated error column name is "Y_error," the direct effect predictor column is named "X_with_error", and its associated errors is "X_error". Finally, we give the heloper function the number of predictor traits, 1 here.

```{r blouch.direct.prep}
############################################################################################################
#Test Blouch prep code - direct effect model - blouch.direct.prep()
dat<-blouch.direct.prep(trdata,"Y_with_error","Y_error","X_with_error","X_error",Z_direct=Z_direct)

```


## Running models
Below are the priors used for this simulation. See Grabowski (in revision) for more on setting these priors. To change these values requires you to open the Stan function, in this case blouchOU_direct, and manually edit them. Unfortunately there is no way around this at present, but trust me - it will be worth it.

For example, here are the four most important priors for the model below - these values are explored in Grabowski (in revision). Remember, always do prior predictive simulations first - in other words, look at distributions of the values and see if they are actually biologically possible - see the exploring priors step above.

```{r, eval=FALSE}
########################################################################################################
#Priors
#hl ~ lognormal(log(0.25),0.25);
#vy ~ exponential(20);
#optima_bar ~ normal(2.0,0,75);
#beta_bar ~ normal(0.25,1.5);
```

And here are the lines of Stan code for setting these priors. To change the values to make them appropriate for your own analyses, you just need to change the numbers below. All Stan programs are in the Blouch/inst/stan folder and named according to the model they run. See Table S1 of Grabowski (in revision) for more on the models. Remember, your priors should be based on what you know about the biological processes underlying your research question and prior predictive simulations (see McElreath 2020)


```{r, eval=FALSE}
#Stan Code
  target += lognormal_lpdf(hl|log(0.25),0.25);
  target += exponential_lpdf(vy|10);
  target += normal_lpdf(alpha|2.0,0.75);
  target += normal_lpdf(beta|0.25,1.5);

```


Now let's run the direct effect model blouchOU_direct.

```{r Run_direct_effect_model , results='hide'}
fit.direct<- rstan::sampling(object = blouch:::stanmodels$blouchOU_direct,data = dat,chains = 2,iter =2000,cores=2)
```


Now let's look at the results
```{r Print_DEM_results}
print(fit.direct,pars = c("hl","vy","alpha","beta"))
#plot(rethinking::precis(fit.direct,depth=3,pars = c("hl","vy","alpha","beta"))) #For use with rethinking package
post<-rstan::extract(fit.direct) #Extract posterior distribution

```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.

```{r}
rm(list=ls())
```

## Adaptive Model
*Blouch* also implements the model of adaptive evolution introduced by Hansen et al. (2008). Here the response variable evolves according to an Ornstein-Uhlenbeck process towards an optimal state that is modeled as a function of the predictor variable.

Again, we will create a phylogeny by randomly sampling from the 10K Trees phylogeny
```{r}
########################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label


```

First we will simulate X and Y data using a generative model for the adaptive model. 

The half-life will be set to 0.1 or 10% of tree height, with the stationary variance (Vy) at 0.01. We will then use the fastBM function from phytools (Revell, 2011) to simulate the X variable following a BM process.

```{r Adaptive_model_setup}
########################################################################################################
#Adaptive Model
########################################################################################################
#Setup parameters
Z_adaptve<-1 #Number of traits
hl<-0.1
a<-log(2)/hl
vy<-0.1 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));
vX0<-0
vY0 <- 0
Sxx<-10 #Look at effects

X<-phytools::fastBM(phy,a=vX0,sig2=Sxx,internal=FALSE) #Simulate X BM variable on tree, with scaling 10
phytools::phenogram(phy,X,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
sigma2_x<-matrix(1,1,1)
```

We will set the intercept (alpha term) to 2 and the slope to 0.25. We will use the *Blouch* helper functions calc_adaptive_dmX to calculate the design matrix and calc_adaptive_V to calculate the variance/covariance matrix.

Then we will simulate mu before using it and V to simulate Y. Finally we make a quick plot of the results.

```{r}
alpha<-2 #Intecept
beta<-0.25 #Slope

dmX<-calc_adaptive_dmX(phy,a,X) #Calculate the design matrix
V<-calc_adaptive_V(phy,a, sigma2_y, beta,  sigma2_x, Z_adaptive) #Calculate adaptive variance/covariance matrix

mu<-alpha+dmX%*%beta #Simulate mu for Y
Y<-MASS::mvrnorm(n=1,mu,V) #Simulate adaptive model Y trait

df<-data.frame(Y=Y,X=X)
names(df)<-c("Y","X")

ggplot2::ggplot(data=df,ggplot2::aes(x=X,y=Y))+
  ggplot2::geom_point()

summary(lm(Y~X,df))

```

Next we will simulate measurement error - we will use a standard deviation of measurement error of 0.01, which we will provide to *Blouch* as a vector (X_error and Y_error), and use the rnorm function to add error to our X and Y variables. In other words, we are telling *Blouch* that the estimated error on X and Y is 0.01, and providing it with X and Y variables that are offset by a random amount of error with this standard deviation.

```{r}
########################################################################################################
#Simulate errors
X_error<-rep(0.01,N)
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-X+rnorm(N,0,0.01)


```

## Data setup for *Blouch*
Next we will use the treeplyr package (Uyeda and Harmon, 2014) make.treedata function to combine the data and tree based on the taxa names. See https://github.com/uyedaj/treeplyr for more on this package.

```{r}
############################################################################################################
#Make trdata file
trait.data<-data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error))
trdata<-treeplyr::make.treedata(phy,trait.data)

############################################################################################################

```

We will use the helper function blouch.adapt.prep() to setup the dat file for Stan. Here the name of the column in trdata$dat that contains the response variable is "Y_with_error", the associated error column name is "Y_error," the direct effect predictor column is named "X_with_error", and its associated errors is "X_error". Finally, we give the helper function the number of predictor traits, 1 here.

```{r blouch.adapt.prep}
#Test Blouch prep code - adaptive model
dat<-blouch.adapt.prep(trdata,"Y_with_error","Y_error","X_with_error","X_error",Z_adaptive=Z_adaptve)

```


## Running models
Below are the priors used for this simulation. See Grabowski (in revision) for more on setting these priors. To change these values requires you to open the Stan function, in this case blouchOU_direct, and manually edit them. Unfortunately there is no way around this at present, but trust me - it will be worth it.

For example, here are the four most important priors for the model below - these values are explored in Grabowski (in revision). Remember, always do prior predictive simulations first - in other words, look at distributions of the values and see if they are actually biologically possible - see the exploring priors step above.

```{r, eval=FALSE}
########################################################################################################
#Priors
#hl ~ lognormal(log(0.25),0.25);
#vy ~ exponential(5);
#alpha ~ normal(2.0,0,2);
#beta_bar ~ normal(0,);
```

And here are the lines of Stan code for setting these priors. To change the values to make them appropriate for your own analyses, you just need to change the numbers below. All Stan programs are in the Blouch/inst/stan folder and named according to the model they run. See Table S1 of Grabowski (in revision) for more on the models. Remember, your priors should be based on what you know about the biological processes underlying your research question and prior predictive simulations (see McElreath 2020)


```{r, eval=FALSE}
#Stan Code
  target += lognormal_lpdf(hl|log(0.25),0.25);
  target += exponential_lpdf(vy|5);
  target += normal_lpdf(alpha|2,0.2);
  target += normal_lpdf(beta|0.25,0);

```


Now let's run the adaptive model blouchOU_adapt.
```{r Run_adaptive_model , results='hide'}
fit.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_adapt,data = dat,chains = 1,iter =400,cores=1)
```

Here the results include both the optimal regression, beta, and the evolutionary regression, beta_e, following Hansen et al. (2008)
```{r Print AM_results}
print(fit.adapt,pars = c("hl","vy","alpha","beta","beta_e"))
#plot(rethinking::precis(fit.adapt,depth=2,pars = c("hl","vy","alpha","beta")))
post<-rstan::extract(fit.adapt)

```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.

```{r}
rm(list=ls())
```

## Direct effect and adaptive models
First we will create a phylogeny by randomly sampling from the 10K Trees phylogeny
```{r}
########################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label


```

Finally we will simulate X and Y data using a generative model for the direct effect and adaptive model. Here we will have 1 direct effect trait, Xd, and one adaptive trait, Xa.

```{r Direct_adapt_model_setup}
############################################################################################################
#Direct effect + Adaptive Model
############################################################################################################
hl<-0.1 #0.1, 0.25, 0.75 - testing options
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));

vX0<-0
vY0 <- 0

Z_direct<-1
Z_adaptive<-1
Z<-Z_direct+Z_adaptive
sigma2_x<-matrix(1,1,1)

Xd<-rnorm(N,0,1)
names(Xd)<-phy$tip.label
phytools::phenogram(phy,Xd,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
Xa<-phytools::fastBM(phy,a=vX0,sig2=sigma2_x[1,1],internal=FALSE) #Simulate X BM variable on tree, with scaling 10
phytools::phenogram(phy,Xa,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
#sigma2_x<-ratematrix(phy,Xa) #Calculate evolutionary v/cv matrix
Xs<-cbind(Xd,Xa)

```

We will have one intercept term here, but two slopes - the order in *Blouch* is always direct effect predictors first, followed by adaptive predictors. Thus the direct effect slope will be 0.35 and the adaptive slope will be 0.25. Except for the *Blouch* helper function calc_mixed_dmX, which is used when there are mixed design matrices, the other functions are the same as the adaptive model above. One complication is the beta term below in the calc_adaptive_V function call - we are sending this function only the adaptive slope and the line beta[(Z_direct+1):(Z_adaptive+Z_direct)] is one way of specifying only that term.

Finally we make a quick plot of the results.

```{r}
alpha<-2 #Intecept
beta<-c(0.35,0.25) #Slopes
dmX<-calc_mixed_dmX(phy,a,Xs,Z_direct,Z_adaptive)
mu<-alpha+dmX%*%beta #Simulate mu for Y

V<-calc_adaptive_V(phy,a, sigma2_y, beta[(Z_direct+1):(Z_adaptive+Z_direct)],  sigma2_x)
Y<-MASS::mvrnorm(n=1,mu,V)

df<-data.frame(Y=Y,X=Xs)

summary(lm(Y~Xs,df))

ggplot2::ggplot(data=df,ggplot2::aes(x=X.Xd,y=Y))+
  ggplot2::geom_point()

summary(lm(Y~X.Xd,df))

ggplot2::ggplot(data=df,ggplot2::aes(x=X.Xa,y=Y))+
  ggplot2::geom_point()

summary(lm(Y~X.Xa,df))

```

Next we will simulate measurement error - we will use a standard deviation of measurement error of 0.01, which we will provide to *Blouch* as a vector (X_error and Y_error), and use the rnorm function to add error to our X and Y variables. In other words, we are telling *Blouch* that the estimated error on X and Y is 0.01, and providing it with X and Y variables that are offset by a random amount of error with this standard deviation. The complication here is we are simulating error in more than one trait - both an adaptive and direct effect trait - the extra lines of code below do this. 

```{r}
########################################################################################################
#Simulate errors - for use with blouchOU_reg_direct_adaptive_ME
Z_X_error<-2 #Number of X traits with error
X_error<-matrix(0.01,nrow=N,ncol=Z_X_error)
X_error<-data.frame(X_error)
names(X_error)<-c("Xd_error","Xa_error")
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-apply(Xs,2,function(X){X+rnorm(N,0,0.01)})

```

## Data setup for *Blouch*
Next we will use the treeplyr package (Uyeda and Harmon, 2014) make.treedata function to combine the data and tree based on the taxa names. See https://github.com/uyedaj/treeplyr for more on this package.

```{r}
############################################################################################################
#Make trdata file
#trdata<-make.treedata(phy,trait.data)
trait.data<-data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error))
trdata<-treeplyr::make.treedata(phy,trait.data)


```

We will use the helper function blouch.direct.adapt.prep() to setup the dat file for Stan. Here the names of the direct effect and adaptive columns are Xd, and Xa and their associated errors, with Z_direct and Z_adaptive the number of direct effect and adaptive traits, respectively.
```{r blouch.direct.adapt.prep}
############################################################################################################
#Test Blouch prep code - Direct effect + Adaptive Model
dat<-blouch.direct.adapt.prep(trdata,"Y_with_error","Y_error",c("Xd","Xa"),c("Xd_error","Xa_error"),Z_direct=1,Z_adaptive=1)

```

Priors are set as above, and the blouchOU_direct_adapt model is run here.

```{r Run_direct_adapt_model , results='hide'}
fit.direct.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_direct_adapt,data = dat,chains = 2,iter =2000,cores=2)

```

Here the results include, in order, the direct effect regression, the optimal regression, and finally beta_e,  the evolutionary regression, following Hansen et al. (2008)

```{r Pring_DAM_results}
print(fit.direct.adapt,pars = c("hl","vy","alpha","beta"))
#plot(rethinking::precis(fit.direct.adapt,depth=2,pars = c("hl","vy","alpha","beta")))
post<-rstan::extract(fit.direct.adapt)


```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.

## References
Revell L.J. 2011. phytools: an R package for phylogenetic comparative biology (and other things). Methods Ecol. Evol. 3:217–223.

Uyeda J.C., Harmon L.J. 2014. A Novel Bayesian Method for Inferring and Interpreting the Dynamics of Adaptive Landscapes from Phylogenetic Comparative Data. Systematic Biology. 63:902–918.

