---
title: "Multi-optima Models-Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-optima Models-Examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
rm(list=ls())
```

```{r setup}
library(blouch)
```



## Multi-optima Model

Next we will run a series of models that include categorical variables - regimes painted on a phylogeny. In all cases the number of iterations is small (400) and we are only running one chain - because this is just an example vignette and these models can take some time to run. In a real analysis you would want to run at least 1000 iterations and at least two chains.




First we will set up the regime painting
```{r Multi_optima_model_setup}
############################################################################################################
#Regimes model
############################################################################################################
########################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label


#Set regimes - manually - 2 regimes
#Locate nodes
plot(phy,no.margin=TRUE,edge.width=2,cex=0.7)
ape::nodelabels(frame="none",adj=c(1.1,-0.4))
ape::tiplabels()

#Paint Regimes on Tree
#source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/blouch/Simulation Code/Functions/set.converge.regimes.R") #Macbook Pro

shifts<-c(84) #Location of nodes with regime shifts
trdata<-data.frame(phy$tip.label)
trdata<-treeplyr::make.treedata(phy,trdata)
trdata<-set.converge.regimes(trdata,shifts)

#Check if manual setting code worked
shifts.total<-c(trdata$dat$regimes,trdata$phy$node.label)
edge.regimes <- factor(shifts.total[trdata$phy$edge[,2]])
print(edge.regimes)

reg.colors <- ggsci::pal_aaas("default", alpha = 0.7)(2)
print(reg.colors)
plot(trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)

```

Format tree 
```{r Format_tree}
############################################################################################################
#Simulate data
n<-length(trdata$phy$tip.label)
regimes_internal <-trdata$phy$node.label
regimes_tip <- trdata$dat$regimes
regimes <- concat.factor(regimes_tip, regimes_internal)
anc_maps<-"regimes"
lineages <- lapply(1:n, function(e) lineage.constructor(trdata$phy, e, anc_maps, regimes)) #Trace lineage from tips (n) to root and determine regimes of each node or branch

```

Now we will simulate Y data based on our generative model
```{r Simulate_Y_data}
############################################################################################################
#Set true values for parameters
hl<-0.1 #0.1, 0.25, 0.75 - testing options
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));
optima<-c(0.5,0.25) #Optima for two regimes

dmX<-weight.matrix(trdata$phy, a, lineages) #Slouch approach
mu<-dmX%*%optima #Simulate mu for Y
V<-calc_direct_V(phy, sigma2_y, a)
Y<-MASS::mvrnorm(n=1,mu,V)

########################################################################################################
#Simulate errors
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)

phytools::phenogram(phy,Y_with_error,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data

############################################################################################################
#Make trdata file
trait.data<-data.frame(cbind(Y_with_error,Y_error))
trdata$dat<-cbind(trdata$dat,data.frame(cbind(Y_with_error,Y_error)))

############################################################################################################
```

We will use the helper function blouch.reg.prep() to setup the dat file for Stan. Here "regimes" is the name of the regime column in trdata$dat.
```{r blouch.reg.prep}
dat<-blouch.reg.prep(trdata,"Y_with_error","Y_error","regimes")
```

Here we run the basic Multi-optima Model, look at the results, and rstan::extract the posterior
```{r Run_RM , results='hide'}
fit.reg<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg,data = dat,chains = 1,iter =400,cores=1)
```

```{r}
print(fit.reg,pars = c("hl","vy","optima"))
##plot(rethinking::precis(fit.reg,depth=2,pars = c("hl","vy","optima")))

post<-rstan::extract(fit.reg) #rstan::extract posterior
```

Next is the code for the Multilevel Multi-optima Model
```{r}
fit.mlm.vi.regimes<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_mlm_vi,data = dat,chains = 2,iter= 2000,cores=2)
print(fit.mlm.vi.regimes,pars = c("hl","vy","optima","optima_bar","sigma"))
###plot(rethinking::precis(fit.mlm.vi.regimes,depth=2,pars = c("hl","vy","optima","optima_bar","sigma")))

post<-rstan::extract(fit.mlm.vi.regimes)
```

Finally we will run the non-centered version of the same model, which can be used when the posterior is hard to explore and the centered version of the model produces divergences.
```{r}
fit.mlm.vi.regimes.nc<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_mlm_vi_nc,data = dat,chains = 1,iter =400,cores=1) #,control=list(adapt_delta=0.95))
print(fit.mlm.vi.regimes.nc,pars = c("hl","vy","optima","optima_bar","sigma"))
###plot(rethinking::precis(fit.mlm.vi.regimes.nc,depth=2,pars = c("hl","vy","optima","optima_bar","sigma")))
post<-rstan::extract(fit.mlm.vi.regimes.nc)
```

```{r}
rm(list=ls())
```

## Multi-optima Direct Effect Model
We will now include both regimes painted on the tree and a direct effect predictor
```{r}
############################################################################################################
#Regimes + Direct Effect Model
############################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label

#Set regimes - manually - 2 regimes
#Locate nodes
plot(phy,no.margin=TRUE,edge.width=2,cex=0.7)
ape::nodelabels(frame="none",adj=c(1.1,-0.4))
ape::tiplabels()

#Paint Regimes on Tree
#source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/blouch/Simulation Code/Functions/set.converge.regimes.R") #Macbook Pro
shifts<-c(84) #Location of nodes with regime shifts
trdata<-data.frame(phy$tip.label)
trdata<-treeplyr::make.treedata(phy,trdata)
trdata<-set.converge.regimes(trdata,shifts)


#Check if manual setting code worked
shifts.total<-c(trdata$dat$regimes,trdata$phy$node.label)
edge.regimes <- factor(shifts.total[trdata$phy$edge[,2]])
print(edge.regimes)

reg.colors <- ggsci::pal_aaas("default", alpha = 0.7)(2)
print(reg.colors)
plot(trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)

#Phylogeny info
n<-length(trdata$phy$tip.label)
mrca1 <- ape::mrca(trdata$phy)
times <- ape::node.depth.edgelength(trdata$phy)

regimes_internal <-trdata$phy$node.label
regimes_tip <- trdata$dat$regimes
regimes <- concat.factor(regimes_tip, regimes_internal)
anc_maps<-"regimes"
lineages <- lapply(1:n, function(e) lineage.constructor(trdata$phy, e, anc_maps, regimes)) #Trace lineage from tips (n) to root and determine regimes of each node or branch

```

Now simulate X and Y data
```{r}
#########################
hl<-0.1 #0.1, 0.25, 0.75 - testing options
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));

vX0<-0
vY0 <- 0
Sxx<-10 #Look at effects

Z_direct<-1

V<-calc_direct_V(phy,sigma2_y,a)
X<-phytools::fastBM(phy,a=vX0,sig2=Sxx,internal=FALSE) #Simulate X BM variable on tree, with scaling 10
#X<-rnorm(N,0,1)
names(X)<-phy$tip.label

phytools::phenogram(phy,X,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data

dmX<-weight.matrix(trdata$phy, a, lineages) #Slouch approach
dmX<-cbind(dmX,X)

beta<-c(2,1,0.25) #Two Optima/One Slope
mu<-dmX%*%beta #Simulate mu for Y

V<-calc_direct_V(phy,sigma2_y,a)
Y<-MASS::mvrnorm(n=1,mu,V)

#Plot data
df<-data.frame(Y=Y,X=X)

ggplot2::ggplot(data=df,ggplot2::aes(x=X,y=Y))+
  ggplot2::geom_point()
summary(lm(Y~X,df))

#################################################################################################################Simulate errors
Z_X_error<-1 #Number of X traits with error
X_error<-rep(0.01,N)
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-X+rnorm(N,0,0.01)

############################################################################################################
#Make trdata file
trdata$dat<-cbind(trdata$dat,data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error)))
```

We will use the helper function blouch.reg.direct.prep() to setup the dat file for Stan. Here "Z_direct" is the number of predictors, and "regimes" is the name of the regime column in trdata$dat.
```{r blouch.reg.direct.prep}
############################################################################################################
#Test Blouch prep code - Regimes + Direct Efffect model
dat<-blouch.reg.direct.prep(trdata,"Y_with_error","Y_error","X_with_error","X_error",Z_direct=1,"regimes")

```

Now we run the basic Multi-optima Direct Effect Model, look at the results, and rstan::extract the posterior
```{r Run_MODEM , results='hide'}
fit.reg.direct<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct,data = dat,chains = 1,iter =400,cores=1)

```

```{r Results_MODEM}
print(fit.reg.direct,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.reg.direct,depth=2,pars = c("hl","vy","optima","beta")))
post<-rstan::extract(fit.reg.direct) #rstan::extract posterior distribution
```

Here is the Multilevel version of the same model
```{r Run_MLMODEM , results='hide'}
fit.mli.regi.direct<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct_mlm_vi,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMODEM}
print(fit.mli.regi.direct,pars = c("hl","vy","optima","optima_bar","beta","sigma"))
##plot(rethinking::precis(fit.mli.regi.direct,depth=2,pars = c("hl","vy","optima","optima_bar","sigma")))
post<-rstan::extract(fit.mli.regi.direct) #rstan::extract posterior distribution

```

And finally the non-centered version of the Multilevel model
```{r Run_MLMODEMNC , results='hide'}
fit.mli.regi.direct<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct_mlm_vi_nc,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMODEMNC}
print(fit.mli.regi.direct,pars = c("hl","vy","optima","optima_bar","beta","sigma"))
###plot(rethinking::precis(fit.mli.regi.direct,depth=2,pars = c("hl","vy","optima","optima_bar","sigma")))
post<-rstan::extract(fit.mli.regi.direct) #rstan::extract posterior distribution

```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.


```{r}
rm(list=ls())
```

## Multi-optima Adaptive Model
We will now include both regimes painted on the tree and an adaptive predictor
```{r Multi_optima_adaptive_model_setup}
############################################################################################################
#Regimes + Adaptive Model
############################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label

#Set regimes - manually - 2 regimes
#Locate nodes
plot(phy,no.margin=TRUE,edge.width=2,cex=0.7)
ape::nodelabels(frame="none",adj=c(1.1,-0.4))
ape::tiplabels()

#Paint Regimes on Tree
shifts<-c(84) #Location of nodes with regime shifts
trdata<-data.frame(phy$tip.label)
trdata<-treeplyr::make.treedata(phy,trdata)
trdata<-set.converge.regimes(trdata,shifts)

#Check if manual setting code worked
shifts.total<-c(trdata$dat$regimes,trdata$phy$node.label)
edge.regimes <- factor(shifts.total[trdata$phy$edge[,2]])
print(edge.regimes)

reg.colors <- ggsci::pal_aaas("default", alpha = 0.7)(2)
print(reg.colors)
plot(trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)

#Phylogeny info
n<-length(trdata$phy$tip.label)

regimes_internal <-trdata$phy$node.label
regimes_tip <- trdata$dat$regimes
regimes <- concat.factor(regimes_tip, regimes_internal)
anc_maps<-"regimes"
lineages <- lapply(1:n, function(e) lineage.constructor(trdata$phy, e, anc_maps, regimes)) #Trace lineage from tips (n) to root and determine regimes of each node or branch


```

Simulate X and Y data using generative model
```{r Multi_optima_adaptive_model_data}
############################################################################################################
#########################
hl<-0.1 #0.1, 0.25, 0.75 - testing options
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));

vX0<-0
vY0 <- 0

Z_adaptive<-1
sigma2_x<-matrix(1,1,1) #Variance of BM Process

X<-phytools::fastBM(phy,a=vX0,sig2=sigma2_x[1,1],internal=FALSE) #Simulate X BM variable on tree, with scaling 10
phytools::phenogram(phy,X,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
#sigma2_x<-ratematrix(phy,Xa) #Calculate evolutionary v/cv matrix

dmX<-weight.matrix(trdata$phy, a, lineages) #Slouch approach
dmX<-cbind(dmX,calc_adaptive_dmX(phy,a,X))

beta<-c(2,1,0.25) #Two Optima/Two Slopes
mu<-dmX%*%beta #Simulate mu for Y

V<-calc_adaptive_V(phy,a, sigma2_y, beta[3],  sigma2_x)
Y<-MASS::mvrnorm(n=1,mu,V)

################################################################################################################
#Plot data
df<-data.frame(Y=Y,X=X)

ggplot2::ggplot(data=df,ggplot2::aes(x=X,y=Y))+
  ggplot2::geom_point()
summary(lm(Y~X,df))
################################################################################################################
#Simulate errors
Z_X_error<-1 #Number of X traits with error
X_error<-rep(0.01,N)
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-X+rnorm(N,0,0.01)

############################################################################################################
#Make trdata file
trdata$dat<-cbind(trdata$dat,data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error)))
############################################################################################################
```

We will use the helper function blouch.reg.adapt.prep() to setup the dat file for Stan. Here "Z_adapt" is the number of predictors, and "regimes" is the name of the regime column in trdata$dat.
```{r blouch.reg.adapt.prep}
dat<-blouch.reg.adapt.prep(trdata,"Y_with_error","Y_error","X_with_error","X_error",Z_adaptive=1,"regimes")

```

Now we run the basic Multi-optima Adaptive Model, look at the results, and rstan::extract the posterior
```{r Run_MOAM , results='hide'}
fit.reg.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_adapt,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MOAM}
print(fit.reg.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.reg.adapt,depth=2,pars = c("hl","vy","optima","beta")))
post<-rstan::extract(fit.reg.adapt) #rstan::extract Posterior Distribution
```

Here is the Multilevel version of the same model
```{r Run_MLMOAM , results='hide'}
fit.reg.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_adapt_mlm_vi,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMOAM}
print(fit.reg.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.reg.adapt,depth=2,pars = c("hl","vy","optima","beta")))
post<-rstan::extract(fit.reg.adapt)
```

And the non-centered version
```{r Run_MLMOAMNC}
fit.reg.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_adapt_mlm_vi_nc,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMOAMNC}
print(fit.reg.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.reg.adapt,depth=2,pars = c("hl","vy","optima","beta")))
post<-rstan::extract(fit.reg.adapt)
```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.


```{r}
rm(list=ls())
```

## Multi-optima Direct Effect and Adaptive Model
We will now include both regimes painted on the tree and a direct effect and adaptive predictor
```{r Multi_optima_direct_adaptive_model_setup}
############################################################################################################
#Regimes + Adaptive + Direct Model
############################################################################################################
#Create phylogeny
########################################################################################################
N<-50 #Number of species
set.seed(10) #Set seed to get same random species each time

phy <- ape::keep.tip(tree.10K,sample(tree.10K$tip.label)[1:N]) 
phy<-ape::multi2di(phy)

l.tree<-max(ape::branching.times(phy)) ## rescale tree to height 1
phy$edge.length<-phy$edge.length/l.tree 

tip.label<-phy$tip.label

#Set regimes - manually - 2 regimes
#Locate nodes
plot(phy,no.margin=TRUE,edge.width=2,cex=0.7)
ape::nodelabels(frame="none",adj=c(1.1,-0.4))
ape::tiplabels()

#Paint Regimes on Tree
shifts<-c(84) #Location of nodes with regime shifts
trdata<-data.frame(phy$tip.label)
trdata<-treeplyr::make.treedata(phy,trdata)
trdata<-set.converge.regimes(trdata,shifts)

#Check if manual setting code worked
shifts.total<-c(trdata$dat$regimes,trdata$phy$node.label)
edge.regimes <- factor(shifts.total[trdata$phy$edge[,2]])
print(edge.regimes)

reg.colors <- ggsci::pal_aaas("default", alpha = 0.7)(2)
print(reg.colors)
plot(trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)

#Phylogeny info
n<-length(trdata$phy$tip.label)

regimes_internal <-trdata$phy$node.label
regimes_tip <- trdata$dat$regimes
regimes <- concat.factor(regimes_tip, regimes_internal)
anc_maps<-"regimes"
lineages <- lapply(1:n, function(e) lineage.constructor(trdata$phy, e, anc_maps, regimes)) #Trace lineage from tips (n) to root and determine regimes of each node or branch

```

```{r Multi_optima_direct_adaptive_model_data}
############################################################################################################
hl<-0.1 #0.1, 0.25, 0.75 - testing options
a<-log(2)/hl
vy<-0.01 #0.25,0.5 - testing options
sigma2_y<-vy*(2*(log(2)/hl));

vX0<-0
vY0 <- 0

Z_direct<-1
Z_adaptive<-1
Z<-Z_direct+Z_adaptive
sigma2_x<-matrix(1,1,1) #Variance of BM Process

Xd<-rnorm(N,0,1)
names(Xd)<-phy$tip.label
phytools::phenogram(phy,Xd,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
Xa<-phytools::fastBM(phy,a=vX0,sig2=sigma2_x[1,1],internal=FALSE) #Simulate X BM variable on tree, with scaling 10
phytools::phenogram(phy,Xa,spread.labels=TRUE,spread.cost=c(1,0)) #Plot X data
#sigma2_x<-ratematrix(phy,Xa) #Calculate evolutionary v/cv matrix
Xs<-cbind(Xd,Xa)

dmX<-weight.matrix(trdata$phy, a, lineages) #Slouch approach
dmX<-cbind(dmX,calc_mixed_dmX(phy,a,Xs,Z_direct,Z_adaptive))

beta<-c(2,1,0.35,0.25) #Two Optima/Two Slopes
mu<-dmX%*%beta #Simulate mu for Y

V<-calc_adaptive_V(phy,a, sigma2_y,   beta[length(beta)],  sigma2_x)
Y<-MASS::mvrnorm(n=1,mu,V)

################################################################################################################
#Plot data
df<-data.frame(Y=Y,Xd=Xs[,1],Xa=Xs[,2])

ggplot2::ggplot(data=df,ggplot2::aes(x=Xd,y=Y))+
  ggplot2::geom_point()

ggplot2::ggplot(data=df,ggplot2::aes(x=Xa,y=Y))+
  ggplot2::geom_point()
summary(lm(Y~Xs,df))

########################################################################################################
#Simulate errors - for use with blouchOU_reg_direct_adaptive_ME
Z_X_error<-2 #Number of X traits with error
X_error<-matrix(0.01,nrow=N,ncol=Z_X_error)
X_error<-data.frame(X_error)
names(X_error)<-c("Xd_error","Xa_error")
Y_error<-rep(0.01,N)
Y_with_error<-Y+rnorm(N,0,0.01)
X_with_error<-apply(Xs,2,function(X){X+rnorm(N,0,0.01)})

############################################################################################################
#Make trdata file
trdata$dat<-cbind(trdata$dat,data.frame(cbind(Y_with_error,Y_error,X_with_error,X_error)))
```

We will use the helper function blouch.reg.direct.adapt.prep() to setup the dat file for Stan. Here the names of the direct effect and adaptive columns are "Xd", and "Xa" and their associated errors, with Z_direct and Z_adaptive th number of direct and adaptive traits, respectively, and "regimes" is the name of the regimes column data in trdata$dat.
```{r blouch.reg.direct.adapt.prep}
############################################################################################################
#Test Blouch prep code - Regimes + Direct Effect + Adaptive Model
dat<-blouch.reg.direct.adapt.prep(trdata,"Y_with_error","Y_error",c("Xd","Xa"),c("Xd_error","Xa_error"),Z_direct=1,Z_adaptive=1,"regimes")

```

First we run the basic Multi-optima Direct Effect Adaptive Model.
```{r Run_MODAM , results='hide'}
fit.reg.direct.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct_adapt,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MODAM}
print(fit.reg.direct.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.adapt,depth=2,pars = c("hl","vy","optima","optima_bar","beta","sigma")))
post<-rstan::extract(fit.reg.direct.adapt) #rstan::extract posterior
```

Next we run the Multilevel version of the same model
```{r Run_MLMODAM , results='hide'}
fit.reg.direct.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct_adapt_mlm_vi,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMODAM}
print(fit.reg.direct.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.adapt,depth=2,pars = c("hl","vy","optima","optima_bar","beta","sigma")))
post<-rstan::extract(fit.reg.direct.adapt)

```

Finally we run the non-centered version of that model
```{r Run_MLMODAMNC , results='hide'}
fit.reg.direct.adapt<- rstan::sampling(object = blouch:::stanmodels$blouchOU_reg_direct_adapt_mlm_vi_nc,data = dat,chains = 1,iter =400,cores=1)
```

```{r Print_MLMODAMNC}
print(fit.reg.direct.adapt,pars = c("hl","vy","optima","beta"))
##plot(rethinking::precis(fit.adapt,depth=2,pars = c("hl","vy","optima","optima_bar","beta","sigma")))

post<-rstan::extract(fit.reg.direct.adapt)

```

From there the posteriors should be explored, compared to the priors, etc. See the Simulation Example for one example of how to do this.

